Efficient and Scalable Bayesian Bipartite Matching with Fast Beta Linkage (fabl)

### 2021-12-20 methods_paper/fabl-paper-blind.tex
updated BA submission

\title{Efficient and Scalable Bayesian Bipartite Matching with Fast Beta Linkage  (fabl)}
\author{Brian Kundinger\textsuperscript{a} \and
  Jerome Reiter\textsuperscript{a} \and 
  Rebecca C.~Steorts \textsuperscript{b}}
\date{
 \textsuperscript{a} Department of Statistical Science, Duke University \\
 \textsuperscript{b} Department of Statistical Science and Computer Science, Duke University\\Principal Mathematical Statistician, United States Census Bureau\\[2ex]
  \today}

@@

We emphasize the computational gains of this split sampler: the first step is a sample from $P$ options, where $P$ does not scale with the size of the linkage task; and the second step is sampling uniformly at random, which is computationally simple even for large sets of candidate records. These changes drastically improve the speed of the sampler, and each can be parallelized if desired for additional computational gains. We provide a summary of the \texttt{fabl} method through pseudocode in Appendix \ref{app:fabl-psuedocode}. {\color{blue}We note here that this split sampler can be adapted for use with \texttt{BRL} as well, but would require updating $\{H_j\}$ to account for the previously matched records that no are longer candidate matches for the current record. \texttt{BRL} would see considerable speed up through the procedure, but this step would keep the complexity of the sampler at $O(n_A \times n_B)$.}

>>>>

\title{Efficient and Scalable Bayesian Bipartite Matching with Fast Beta Linkage  (fabl)}
%\author{Brian Kundinger\textsuperscript{a} \and
%  Jerome Reiter\textsuperscript{a} \and 
%  Rebecca C.~Steorts \textsuperscript{b}}
%\date{
% \textsuperscript{a} Department of Statistical Science, Duke University \\
% \textsuperscript{b} Department of Statistical Science and Computer Science, Duke University\\Principal Mathematical Statistician, United States Census Bureau\\[2ex]
%  \today}

@@

We emphasize the computational gains of this split sampler: the first step is a sample from $P$ options, where $P$ does not scale with the size of the linkage task; and the second step is sampling uniformly at random, which is computationally simple even for large sets of candidate records. These changes drastically improve the speed of the sampler, and each can be parallelized if desired for additional computational gains. We provide a summary of the \texttt{fabl} method through pseudocode in Appendix \ref{app:fabl-psuedocode}. We note here that this split sampler can be adapted for use with \texttt{BRL} as well, but would require updating $\{H_j\}$ to account for the previously matched records that no are longer candidate matches for the current record. \texttt{BRL} would see considerable speed up through the procedure, but computation time would still scale with the size of the larger dataset.
